Renewal and stability in populations structured by remaining years of life

### 2014-09-11 ThanoRepro/ThanoRepro.tex
changes to perron frob paragraph

For rate schedules typical of human populations, thanatological projection
matrices are irreducible and primitive, and so will always have a
dominant positive real eigenvalue, the natural log of which is $r$. This
observation is sufficient to demonstrate the strong ergodic
theorem:\footnote{This step was taken for granted in jumping from
equation~\eqref{eq:ergostep1} to ~\eqref{eq:ergostep2}} i.e., given a fixed
matrix and enough time, the population will conform to some stable thanatological age distribution. The eigenvector corresponding to the dominant eigenvalue gives this stable thanatological age structure. The square
of the, or the pace of convergence to stability (the ratio of the \nth{1} to the \nth{2} eigenvalues).
\footnote{See \citet[p.86-87]{caswell2001matrix}.}

>>>>

For rate schedules typical of human populations, thanatological projection
matrices are filled with non-negative values, mostly greater than zero. Raising
the matrix to some not-very-large power, $k$, will make all entries greater than zero. This observation, or else by noting that the
lifecylce graph is strongly connected, lets us conclude that the matrix is
irreducible and primitive. By the Perron-Frobenius theorem, the matrix will
 always have a unique, dominant, positive real eigenvalue, the natural
log of which is the intrinsic growth rate, $r$, and strong ergodicity is
assured.
In other words, given a fixed matrix and enough time, the population will conform to some stable thanatological age distribution. Strong
ergodicity was taken for granted in jumping from equation~\eqref{eq:ergostep1}
to ~\eqref{eq:ergostep2}, but the presently described properties of the matrix
model should, albeit unrigorously, satisfy lingering uncertainty on that step.
The eigenvector corresponding to the dominant eigenvalue gives this stable
thanatological age structure. If the fertility rates placed in the matrix are
from the stable population, then the growth rate from the thanatological
projection matrix is equal to that of the standard Leslie
matrix.\footnote{Some care needs to be taken in order to demonstrate this
point, since common approximations and adjustments used in matrix
construction can throw things off. \texttt{R} code is available from the
author to demonstrate this point. See also Appendix~\ref{app:B} for the
corresponding proof from the continuous model.} The square of the, or the pace
of convergence to stability (the ratio of the \nth{1} to the \nth{2} eigenvalues).
\footnote{See \citet[p.86-87]{caswell2001matrix}.}
